03-computation_issues

### 2019-02-08 journals/03-computation_issues.Rmd
worked more on reorganizing journals

---
title: "Applying LIME to Bullet Data Research Journal"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE, 
                      fig.pos = "center")
```

@@

I decided to start a journal to keep track of the work I am doing in regards to applying LIME to the Hamby bullet data. I want a place to store my work to use for future purposes and where I do not feel the need to "perfect" code. I am hoping that this will be a place where I can write drafts of code and create first versions of graphics. Then I can clean up and organize the code when I write the papers.

Prior to this journal, I had been working on all of my ideas in the R markdown document for the firearm examiner paper, but now I am going to transfer a lot of that code to this journal. I am transferring them here now. Later I will move the necessary parts to the paper.

```{r}
# Load libraries
library(bulletr)
library(caret)
library(cowplot)
library(furrr)
library(future)
library(gretchenalbrecht)
library(irr)
library(lime)
library(plotly)
library(purrr)
library(randomForest)
library(tidyverse)

# Source functions
source("../code/helper_functions.R")

# Load in the training data (Hamby Data 173 and 252) and testing data (Hamby Data 224 Sets 1 and 11)
hamby173and252_train <- read.csv("../data/hamby173and252_train.csv")
hamby224_test <- read.csv("../data/hamby224_test.csv") # created in the paper

# Read in the test_explain data
hamby224_test_explain <- readRDS("../data/hamby224_test_explain.rds")

# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(rtrees$importance)
```

# Computation Issues

## Plotly Dead Space Issue

@@

## Caret vs RandomForest

@@

## Comparing Furrr Times

@@

# New Binning Methods

## Subsampled Bins

One idea that I had to try to improve the lime explanations was to subsample from the training dataset before computing the quantiles. The proportion of observations with `samesource = FALSE` is very large in the training dataset, so that overwhelms the observations with `samesource = TRUE`. By subsampling, it might allow for better locations to bin the data. I wrote the code below to attempt to apply lime to a subsampled version of the training dataset.

```{r eval = FALSE}
# Code saved for future use for the sampling data

# Create the data (with a set seed)
set.seed(20181128)
hamby173and252_train_sub <- hamby173and252_train %>%
  filter(samesource == FALSE) %>%
  slice(sample(1:length(hamby173and252_train$barrel1), 4500, replace = FALSE)) %>%
  bind_rows(hamby173and252_train %>% filter(samesource == TRUE))

# Save the subsampled data
write.csv(hamby173and252_train_sub, "../data/hamby173and252_train_sub.csv", row.names = FALSE)

# Apply lime to the subsampled training data with the specified input options
hamby224_lime_explain_sub <- future_pmap(.l = as.list(hamby224_lime_inputs %>%
                                                        select(-case)),
                                         .f = run_lime, # run_lime is one of my helper functions
                                         train = hamby173and252_train_sub %>% select(rf_features),
                                         test = hamby224_test %>% arrange(case) %>% select(rf_features) %>% na.omit(),
                                         rfmodel = as_classifier(rtrees),
                                         label = "TRUE",
                                         nfeatures = 3,
                                         seed = TRUE) %>%
  mutate(training_data = factor("Subsampled"))

# Separate the lime and explain function results from the subsampled data
hamby224_lime_sub <- map(hamby224_lime_explain_sub, function(list) list$lime)
hamby224_explain_sub <- map_df(hamby224_lime_explain_sub, function(list) list$explain)

# Join the lime results from the full and subsampled training data
hamby224_lime <- c(hamby224_lime_full, hamby224_lime_sub)
hamby224_explain <- bind_rows(hamby224_explain_full, hamby224_explain_sub)
```

## Tree Based Bins

Heike suggested the idea of using classification trees to choose the cutoffs for the lime bins. That is, the divisions in the tree could be used as the bin cuts. This would allow us to automate the process and to obtain a penalty parameter since the trees give us nesting. (MSE + lambda * p where p is the from the number of trees multiplied by the number of ...)

Heike wrote the function `treebink` to take x and y variables for fitting a tree and returning the cuts for $k$ bins based on the splits from the tree. The function is stored in the file   `helper_functions.R`. Here is an example using the function below with the predictor variable of ccf.

```{r}
# Example using treebink
treebink(y = hamby173and252_train$samesource, x = hamby173and252_train$ccf, k = 5)
```

However, there seems to be a problem with `treebink`. For certain variables, when a high number of bins is requested, the function returns more than the desired number of bins. Heike is going to look into this.

She recently updated the function `treebink` to run using the tree package. Here are some comments from her about the code:

"I've moved it from rpart to tree. I am not quite sure that the two packages are doing the exact same thing. They claim they do but they might be implemented a bit differently. What we are doing now is if there are too many splits, we look at the order in which the splits are made and roll back some of them. I haven't found a case that doesn't work, but obviously that doesn't show that there isn't any :)"

I wrote the function `mylime`, which adds an option in the `lime` function from the LIME package to use Heike's function `treebink` to obtain tree based bins. This function is also included in the file `helper_functions.R`. The output from `mylime` can then be input into the `explain` function to obtain explanations. I was able to rerun all of the code for creating the `hamby224_test_explain` dataset with the explanations from the tree based bins included for 2 to 6 bins.

# Assessing Lime Explanations

When we began looking at the explanations from `lime` with the default settings, we did not think that they made sense. This led me to try applying LIME with a handful of input values. I also have tried running LIME with the tree based bins. Additionally, since LIME is based on random permutations, I was curious to know how consistent the results were. This led me to try running each implementation a handful of times. I have attempted to evaluate these results by both considering the accuracy and consistency of the results.

## Accuracy

The figures in this section are created from the implementations of LIME on set 1 from the training data for the input options of 2 to 6 quantile bins, 2 to 6 equally spaced bins, 2 to 6 tree based bins, kernel density estimation, and normal distribution approximation. Each set of input values was only run once.

```{r}
# Read in the lime comparison data (the comparisons are run in the firearm examiner
# paper since they will be discussed in the paper)
hamby224_lime_comparisons <- readRDS("../data/hamby224_lime_comparisons.rds")
```

### Complex versus Simple Model Predictions

The plot below compares the predictions from the "simple model" (the ridge regression model) and the "complex model" (the random forest `rtrees`) on the x-axis from the lime implementations with bin estimation (`bin_continuous == TRUE`). The simple model predictions are on the y-axis, and the complex model predictions are on the x-axis. The plot is faceted by number of bins and the type of bin. The points are colored by the $R^2$ value from the fit of the simple model. The lines are linear regression lines fit to the data points within a facet. We would hope that there is a linear relationship between these two variables. None of the cases show strong linear trends, but some are more linear than others. The quantile bins show that the simple model never makes a prediction over 0.6, whereas the random forest model can have predictions of up to 1. The equally spaced bins do have probabilities that exceed 0.6, but only with 3 and 6 bins. The tree based bins do the best job with predictions above 0.6 for all number of bins. The tree bins also seem to group the simple model predictions into two categories (probably based on the samesource variable). I noticed that within the facets, the points are in mostly horizontal strips, and the number of strips is about the same as the number of bins from the lime implementation.

```{r fig.height = 8, fig.width = 10}
# Plot of simple model versus complex model predictions
hamby224_lime_comparisons %>%
  filter(!is.na(rfscore), set == "Set 1", bin_continuous == TRUE) %>%
  ggplot(aes(x = rfscore, y = model_prediction)) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(bin_situation ~ nbins) + 
  theme_bw() + 
  stat_smooth(method = "lm", se = FALSE, size = 0.5) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Complex Model Prediction (RF Score)", y = "Simple Model Prediction",
       color = "Simple Model R2")
```

The plot below shows the absolute value of the difference between the complex model prediction and the simple model prediction versus the complex model prediction. The points are faceted by number of bins and bin type. Again, the points are colored by the $R^2$ values from the simple model. All cases show a v-shaped trend. The low part of the v occurs around a random forest score of about 0.25. That is, the simple model is most accurately portraying the complex model predictions around the random forest score of 0.25. It gets worse near the extremes. However, it is interesting to note that with the equally spaced bins and the tree based bins, the absolute difference decreases near 1. This is not the case with the quantile bins. It appears that the equally spaced bins and tree based bins are able to make slightly better predictions than the quantile bins when there is a high probability of a match.

```{r fig.height = 8, fig.width = 10}
# Plot of difference in predictions versus rfscore
hamby224_lime_comparisons %>%
  filter(!is.na(rfscore), set == "Set 1", bin_continuous == TRUE) %>%
  ggplot(aes(x = rfscore, y = abs(diff))) + 
  geom_point(aes(color = model_r2)) + 
  facet_grid(bin_situation ~ nbins) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Complex Model Prediction (RF Score)", 
       y = "Difference Between Complex and Simple Model Predictions",
       color = "Simple Model R2")
```

At some point, I got the idea in my head that it would be interesting to look at the "residuals" (difference in complex and simple model predictions) by feature. The plot below shows one example with the data from 3 equally spaced bins. There are clear trends in the plots, but I am not quite sure what to make of this or how to use it. This may be something to return to.

```{r}
# Plots of model "residuals"
hamby224_lime_comparisons %>%
  filter(quantile_bins == FALSE, bin_method == "equally_spaced", nbins == "3", set == "Set 1") %>%
  select(rf_features, diff, -case) %>%
  gather(key = feature, value = feature_value, rf_features) %>%
  select(feature, feature_value, diff) %>%
  ggplot(aes(x = feature_value, y = diff)) + 
  geom_point() + 
  facet_wrap(~ feature, scales = "free") + 
  geom_hline(yintercept = 0, color = "blue") + 
  labs(x = "Feature Value", y = "Residual")
```

### Comparing Input Values by MSE and $R^2$

In order to assess which lime implementation is doing the best job of capturing the predictions from the random forest model, I decided to calculate the mean squared error for each of the input situations. I defined the mean squared error as 
  $$\frac{\sum_{i=1}^n (\hat{p}_{\mbox{simple},i}-\hat{p}_{\mbox{complex},i})^2}{n}$$
where $n$ is the number of observations in the testing dataset (within a set). Additionally, I was curious to compare the fits of the ridge regression models across input values, so I calculated the average $R^2$ for each input situation.

```{r}
# Summarizing lime results
hamby224_lime_results <- hamby224_lime_comparisons %>%
  filter(!is.na(rfscore)) %>%
  group_by(situation, bin_situation, bin_continuous, quantile_bins, nbins, use_density, 
           bin_method, response, set) %>%
  summarise(mse = (sum(diff^2)) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  arrange(set) %>%
  ungroup()
```

The plot below shows the mean squared errors for each of the input situations faceted by set. In both cases, the mean squared error with 3 equally spaced bins is low, and the MSE for the tree binning cases are also low.

```{r}
# Plot of MSEs
hamby224_lime_results %>%
  mutate(bins = ifelse(nbins == 4 & bin_continuous == FALSE, "other", nbins)) %>%
  ggplot(aes(x = bin_situation, y = mse, color = bin_situation)) +
  geom_point() + 
  facet_grid(set ~ bins, scales = "free_x", space = "free_x") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "bottom") + 
  labs(x = "Bin Type", y = "MSE", color = "") + 
  scale_color_gretchenalbrecht(palette = "winged_spill", discrete = TRUE)
```

The plot below shows the average $R^2$ value for each of the input situations faceted by set. The results are very similar for each set, and the highest $R^2$ values occur for 2 quantile bins. The $R^2$ values for 3 equally spaced bins are in the middle, and the $R^2$ values for all of the tree based bins are the lowest.

```{r}
# Plot of R^2s
hamby224_lime_results %>%
  mutate(bins = ifelse(nbins == 4 & bin_continuous == FALSE, "other", nbins)) %>%
  ggplot(aes(x = bin_situation, y = ave_r2, color = bin_situation)) +
  geom_point() + 
  facet_grid(set ~ bins, scales = "free_x", space = "free_x") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") + 
  labs(x = "Bin Type", y = "Average R2", color = "") + 
  scale_color_gretchenalbrecht(palette = "winged_spill", discrete = TRUE)
```

## Consistency with One Implementation

I wanted to look at the consistency of features chosen by lime across the number of bins within different bin types. I would like the features chosen to not depend on the number of bins used. Otherwise, it will be difficult to know how many bins to use. Additionally, it would be nice if the features chosen differed across cases, so that the explanations are based on the locality of the observation and not general across all locations. The code below creates a dataset that identifies the order in which the features are chosen by lime.

```{r}
# Creates a dataset that creates columns of the chosen features by lime from first to third
chosen_features <- hamby224_test_explain %>%
  filter(!(bin_situation %in% c("kernel density", "normal approximation"))) %>%
  filter(!is.na(rfscore)) %>%
  select(set, situation, bin_situation, nbins, case, samesource, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(situation, case, desc(feature_weight_abs)) %>%
  mutate(explainer = rep(c("first", "second", "third"), length(situation) / 3)) %>%
  select(-feature_weight, -feature_weight_abs) %>%
  spread(explainer, feature)
```

In order to access consistency, I have decided to compute Fleiss' kappa for each of the bin types. Fleiss' kappa is often used to assess the consistency across raters for any number of raters and with nominal ratings. It calculates the degree of agreement in classification over that which would be expected by chance. A description of how it is computed can be found on the wikipedia site: https://en.wikipedia.org/wiki/Fleiss%27_kappa

I computed kappa for each bin situation separately for the top three features chosen. These values are plotted below for each set. For both the first and second features chosen by lime, the values of kappa suggest that quantile bins are the most consistent across all the number of bins followed by the tree based bins. The values of kappa for the quantile bins are slight to fair agreement. For the third chosen variable, the tree based bins have the highest value of kappa, but all the kappa values for the third chosen feature are low.

```{r}
# Computes and plots the values of kappa
chosen_features %>%
  ungroup() %>%
  select(-situation) %>%
  gather(key = chosen, value = feature, first:third) %>%
  spread(nbins, feature) %>%
  select(-case) %>%
  rename(bins2 = "2", bins3 = "3", bins4 = "4", bins5 = "5", bins6 = "6") %>%
  group_by(set, bin_situation, chosen) %>%
  summarise(kappa = 
              irr::kappam.fleiss(matrix(c(bins2, bins3, bins4, bins5, bins6), 
                                        ncol = 5))$value) %>%
  arrange(chosen, set, bin_situation) %>%
  ggplot(aes(x = bin_situation, y = kappa, color = chosen, group = chosen)) +
  geom_point() + 
  geom_line() +
  facet_grid( ~ set) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_color_gretchenalbrecht(palette = "pink_cloud", discrete = TRUE) +
  labs(x = "Bin Type", y = "Kappa", color = "Feature")
```

I also attempted to plot the features chosen by lime in order to visually assess the consistency. This plots are included below.

```{r fig.height = 8, fig.width = 8}
# Plots the top feature for each case for each bin situation
p <- ggplot(chosen_features, aes(x = situation, y = case, fill = first)) + 
  geom_tile() + 
  facet_grid(set + samesource ~ bin_situation, scales = "free", space = "free_y") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) + 
  labs(x = "Bin Type", y = "Case", fill = "Feature", title = "First Feature")

# Plots the second feature for each case for each bin situation
ggplot(chosen_features, aes(x = situation, y = case, fill = second)) + 
  geom_tile() + 
  facet_grid(set + samesource ~ bin_situation, scales = "free", space = "free_y") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) + 
  labs(x = "Bin Type", y = "Case", fill = "Feature", title = "Second Feature")

# Plots the third feature for each case for each bin situation
ggplot(chosen_features, aes(x = situation, y = case, fill = third)) + 
  geom_tile() + 
  facet_grid(set + samesource ~ bin_situation, scales = "free", space = "free_y") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) + 
  labs(x = "Bin Type", y = "Case", fill = "Feature", title = "Third Feature")
```

I also created the plot below to visually assess the proportions of cases within a bin situation and level of feature chosen by lime (first, second, or third) that have 1 to 5 chosen features. The plots are separated by set.

```{r}
# Plots the proportions of number of top features chosen across number of bins for a case
# within a binning method
chosen_features %>%
  group_by(set, case, bin_situation) %>%
  summarise(first = length(levels(factor(first))),
            second = length(levels(factor(second))),
            third = length(levels(factor(third)))) %>%
  gather(chosen, nfeatures, first:third) %>%
  ggplot(aes(x = bin_situation, fill = factor(nfeatures))) + 
  geom_bar(position = "stack") + 
  facet_grid(set ~ chosen) + 
  labs(x = "Bin Type", y = "Count", fill = "Number of Features") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") + 
  scale_fill_gretchenalbrecht(palette = "winter_light", discrete = TRUE)
```

## Consistency Across Implementations

In order to assess how variable the LIME results are, I ran each of the input situations 10 times. I wanted to look at the variability of the MSE for each situation, and I wanted to see how consistent the choice of selected features is by LIME.

```{r}
# Perform the sensitivity analysis if not already saved
if(!file.exists("../data/hamby224_sensitivity_inputs.rds")) {

  # Specify the number of reps and input cases
  nreps = 10
  noptions = 22

  # Specify the inputs for the sensitivity analysis
  hamby224_sensitivity_inputs <- list(bin_continuous = c(rep(TRUE, nreps * 20),
                                                         rep(FALSE, nreps * 2)),
                                      quantile_bins = c(rep(c(TRUE, FALSE), each = nreps * 5), 
                                                        rep(TRUE, nreps * 12)),
                                      nbins = c(rep(rep(2:6, each = nreps), 4), 
                                                rep(4, nreps * 2)),
                                      use_density = c(rep(TRUE, nreps * 21), 
                                                      rep(FALSE, nreps)),
                                      bin_method = c(rep("quantile_bins", nreps * 5),
                                                      rep("equally_spaced", nreps * 5),
                                                      rep("tree", nreps * 10),
                                                      rep("quantile_bins", nreps * 2)),
                                      response = c(rep(NA, nreps * 10),
                                                   rep("samesource", nreps * 5),
                                                   rep("rfscore", nreps * 5),
                                                   rep(NA, nreps * 2)))
  # Tell R to run the upcoming code in parallel
  plan(multiprocess)

  # Run lime for the sensitivity analysis and organize the output
  hamby224_sensitivity_outputs <- future_pmap(.l = hamby224_sensitivity_inputs,
             .f = run_lime, # run_lime is one of my helper functions
             features = rf_features,
             train = hamby173and252_train,
             test = hamby224_test %>% 
               arrange(case) %>% 
               select(rf_features) %>% 
               na.omit(),
             rfmodel = as_classifier(rtrees),
             label = "TRUE",
             nfeatures = 3,
             seed = FALSE) %>%
    map_df(function(list) list$explain) %>%
    mutate(rep = factor(rep(rep(1:nreps, each = dim(hamby224_test %>% na.omit())[1] * 3),
                            noptions)))

  # Turn the input options into a dataframe to be saved
  hamby224_sensitivity_inputs <- hamby224_sensitivity_inputs %>%
    unlist() %>%
    matrix(ncol = length(hamby224_sensitivity_inputs), 
           dimnames = list(NULL, names(hamby224_sensitivity_inputs))) %>%
    as.data.frame() %>%
    mutate(case = 1:(nreps * noptions)) %>%
    select(case, bin_continuous:bin_method)

  # Save the sensitivity inputs and outputs
  saveRDS(hamby224_sensitivity_inputs, "../data/hamby224_sensitivity_inputs.rds")
  saveRDS(hamby224_sensitivity_outputs, "../data/hamby224_sensitivity_outputs.rds")

} else {

  # Load in the sensitivity inputs and outputs
  hamby224_sensitivity_inputs <- readRDS("../data/hamby224_sensitivity_inputs.rds")
  hamby224_sensitivity_outputs <- readRDS("../data/hamby224_sensitivity_outputs.rds")

}
```

```{r}
# Join the sensativity outputs with the test data
hamby224_sensitivity_joined <- hamby224_sensitivity_outputs %>%
  full_join(hamby224_test %>% na.omit() %>% 
                mutate(case = as.character(case)), by = "case") %>%
    mutate(case = factor(case)) %>%
    select(case, model_r2:feature_weight, bin_continuous:rep, 
           set:land2, rfscore, samesource) %>%
    mutate(diff = rfscore - model_prediction,
         nbins = as.numeric(as.character(nbins)),
         situation = ifelse(bin_continuous == TRUE & bin_method == "quantile_bins", 
                            sprintf("%.0f quantile", nbins),
                            ifelse(bin_continuous == TRUE & bin_method == "equally_spaced",
                                   sprintf("%.0f equally spaced", nbins),
                                   ifelse(bin_continuous == TRUE & bin_method == "tree" &
                                            response == "samesource",
                                          sprintf("%.0f samesource tree", nbins),
                                          ifelse(bin_continuous == TRUE & bin_method == "tree" &
                                            response == "rfscore",
                                            sprintf("%.0f rfscore tree", nbins),
                                            ifelse(bin_continuous == FALSE & 
                                                   use_density == TRUE, 
                                                   "kernel density", 
                                                   "normal approximation"))))) %>%
           fct_relevel("2 quantile", "3 quantile", "4 quantile",
                       "5 quantile", "6 quantile", "2 equally spaced",
                       "3 equally spaced", "4 equally spaced",
                       "5 equally spaced", "6 equally spaced",
                       "2 samesource tree", "3 samesource tree",
                       "4 samesource tree", "5 samesource tree",
                       "6 samesource tree")) %>%
    mutate(bin_situation = ifelse(bin_method == "quantile_bins" & 
                                  bin_continuous == TRUE,
                                  "quantile",
                                  ifelse(bin_method == "equally_spaced" & 
                                         bin_continuous == TRUE,
                                         "equally spaced", 
                                         ifelse(bin_method == "tree" & 
                                                bin_continuous == TRUE & 
                                                response == "samesource",
                                                "samesource tree", 
                                                ifelse(bin_method == "tree" & 
                                                       bin_continuous == TRUE & 
                                                       response == "rfscore",
                                                       "rfscore tree", 
                                                       ifelse(bin_continuous == FALSE & 
                                                       use_density == TRUE, 
                                                       "kernel density", 
                                                       "normal approximation")))))) %>%
    mutate(bin_situation = factor(bin_situation))
```

```{r}
# Code for organizing the sensitivity analysis results
hamby224_sensitivity_results <- hamby224_sensitivity_joined %>%
  group_by(set, rep, situation, bin_situation, nbins) %>%
  summarise(mse = sum((diff)^2) / length(diff),
            ave_r2 = mean(model_r2))
```

The plot below shows boxplots of the mean squared errors computed for each of the input situations for each set. There does not appear to be much variation for most of the cases. The largest amount of variation occurs with 4 and 5 equally spaced bins. There is moderate variability for 3 equally spaced bins. The lowest MSEs occur for the samesource tree based bins followed by the rfscore tree based bins for most cases.

```{r}
# Plots the MSEs from the sensativity analysis
hamby224_sensitivity_results %>%
  mutate(bins = ifelse(bin_situation %in% c("kernel density", "normal approximation"),
                       "other", as.character(nbins))) %>%
  ggplot(aes(x = bin_situation, y = mse)) + 
  geom_boxplot(aes(color = bin_situation)) +
  facet_grid(set ~ bins, scales = "free_x", space = "free_x") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
  labs(x = "Bin Type", y = "MSE", color = "") + 
  scale_color_gretchenalbrecht(palette = "winged_spill", discrete = TRUE)
```

The plot blow shows boxplots of the average $R^2$ values from the 10 models for all input options and for each set. There is minimal variability in the average $R^2$ value for all cases, and the relationships between $R^2$ values and input cases agree with the single implementation seen below.

```{r}
# Plots the average R^2 value from the sensativity analysis 
hamby224_sensitivity_results %>%
  mutate(bins = ifelse(bin_situation %in% c("kernel density", "normal approximation"),
                       "other", as.character(nbins))) %>%
  ggplot(aes(x = bin_situation, y = ave_r2)) + 
  geom_boxplot(aes(color = bin_situation)) +
  facet_grid(set ~ bins, scales = "free_x", space = "free_x") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
  labs(x = "Bin Type", y = "MSE", color = "") + 
  scale_color_gretchenalbrecht(palette = "winged_spill", discrete = TRUE)
```

To assess the consistency of the results, I determined the number of different features chosen by LIME as the top feature within a test case across the 10 replicates. I then computed the average number of different top features chosen by LIME within a input situations for each set. The plot below contains these values shown as the dots. The error bars represent one standard deviation above and below the mean. Note that the number of levels cannot go below 1, but the error bars go below 1. I have decided to ignore this for now, but I would need to come up with a better representation if I used this for something.

There are a handful of situations with a mean near one and very little variation. However, the largest mean is around 2, which is not very high. This seems to suggest that LIME is relatively consistent across runs.

```{r}
# Determine and plot the average and variability of the number levels for each case 
# and input option
hamby224_consistency <- hamby224_sensitivity_joined %>%
  select(set, situation, bin_situation, nbins, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(set, situation, case, rep, desc(feature_weight_abs)) %>%
  group_by(set, situation, nbins, case, rep) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(set, situation, bin_situation, nbins, case) %>%
  summarise(nlevels = length(levels(factor(feature))),
            count = n()) %>%
  ungroup() %>%
  group_by(set, situation, bin_situation, nbins) %>%
  summarise(mean_nlevels = mean(nlevels),
            sd_nlevels = sd(nlevels),
            upper = mean(nlevels) + sd(nlevels),
            lower = mean(nlevels) - sd(nlevels),
            max = max(nlevels),
            min = min(nlevels)) %>%
  mutate(bins = ifelse(bin_situation %in% c("kernel density", "normal approximation"), 
                       "other", as.character(nbins)))

hamby224_consistency %>%
  ggplot(aes(x = situation, y = mean_nlevels)) + 
  geom_errorbar(aes(ymin = lower, ymax = upper), color = "grey80") +
  geom_point(aes(color = bin_situation)) + 
  facet_grid(set ~ bins, scales = "free_x", space = "free_x") + 
  labs(x = "Bin Type", y = "Mean Number of Features", color = "") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") + 
  scale_color_gretchenalbrecht(palette = "winged_spill", discrete = TRUE)
```

The bar chart below shows the proportion of different number of first, second, and third features chosen across all test cases within a set for each input situation. This shows that the most different features chosen was seven. However, many of the cases only have one or two different features chosen, and usually, a larger proportion of the cases have only one top feature chosen across the 10 reps.

```{r}
# Determine the order that the features are chosen within a case and 
# set of input options
sensitivity_chosen <- hamby224_sensitivity_joined %>%
  select(set, situation, bin_situation, nbins, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(set, situation, case, rep, desc(feature_weight_abs)) %>%
  mutate(explainer = rep(c("first", "second", "third"), length(situation) / 3)) %>%
  select(-feature_weight, -feature_weight_abs) %>%
  spread(explainer, feature)
```

```{r fig.width = 10}
# Determine and plot the number of features chosen across the 10 reps within 
# a set, input options, and feature level
sensitivity_chosen %>%
  group_by(set, situation, case) %>%
  summarise(first = length(levels(factor(first))),
            second = length(levels(factor(second))),
            third = length(levels(factor(third)))) %>%
  gather(chosen, nfeatures, first:third) %>%
  ggplot(aes(x = situation, fill = factor(nfeatures))) + 
  geom_bar(position = "stack") + 
  facet_grid(set ~ chosen) + 
  labs(x = "Bin Type", y = "Count", fill = "Number of Features") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") + 
  scale_fill_gretchenalbrecht(palette = "flood_tide", discrete = TRUE, reverse = FALSE)
```

I computed Fleiss's kappa for each sampling technique within a set across the 10 reps. These values are plotted below. Right now, the quantile methods and rfscore tree based bins seem to be the most consistent. However, I am not happy with some of the kappa values for cases where the next set of plots would suggest a high kappa, but the value is very low. I wonder if this a flaw in the Fleiss's kappa and would suggest that I use a different metric.

```{r fig.width = 10}
sensitivity_chosen %>%
  mutate(bins = ifelse(bin_situation %in% c("kernel density", "normal approximation"), 
                       "other", 
                       nbins)) %>%
  gather(key = chosen, value = feature, first:third) %>%
  spread(rep, feature) %>%
  rename(rep1 = "1", rep2 = "2", rep3 = "3", rep4 = "4", rep5 = "5", 
         rep6 = "6", rep7 = "7", rep8 = "8", rep9 = "9", rep10 = "10") %>%
  group_by(set, situation, bin_situation, bins, chosen) %>%
  summarise(kappa = 
              irr::kappam.fleiss(matrix(c(rep1, rep2, rep3, rep4, rep5,
                                          rep6, rep7, rep8, rep9, rep10), 
                                        ncol = 10))$value,
            nunique = length(unique(c(rep1, rep2, rep3, rep4, rep5, 
                                      rep6, rep7, rep8, rep9, rep10)))) %>%
  mutate(kappa = ifelse(is.nan(kappa), 1, kappa)) %>%
  ggplot(aes(x = bin_situation, y = kappa, color = chosen, group = chosen)) + 
  geom_point() + 
  geom_line(aes(linetype = chosen)) +
  facet_grid(set ~ bins, scales = "free_x", space = "free_x") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_color_gretchenalbrecht(palette = "pink_cloud", discrete = TRUE) +
  labs(x = "Sampling Technique", y = "Kappa", color = "Feature", linetype = "Feature")
```

The plots below need some work, but they contain some interesting information. Right now, I am only concentrating on set 1. We can see that with some lime settings, all cases have the same "best" variable chosen, and other settings have multiple important variables that depend on the case. We can also see that different variables are selected as important depending on the number of bins. It is also interesting to note that when comparing the quantile bins to the equally spaced bins, this plot shows that the equally spaced bins tend to choose the same first variable for all cases. On the other hand, the quantile bins choose different first variables for the cases.

The fact that different features get chosen as the top feature for different estimation techniques makes sense. You would expect that the number of bins would determine which features are better suited for predicting whether or not a comparison is a match. For example, if you look back at the feature distribution plots, you can see that ccf is the obvious choice for determining between matches and non-matches if you have to choose two equally spaced bins. This is the feature that lime most frequently chooses. You can make similar arguments for other cases as well. Even though different inputs select different variables, I would expect that some variables will be better at prediction than others. I think this is what leads some input values to better LIME explanations and lower MSEs than others.

In the meeting with Heike, we talked about how this plot suggests that LIME is not doing a very good job of providing local explanations. Regardless of the case, it is often suggesting only one of two variables are important, and the variable importance depends on the number of bins. In this plot, I would like consistency in the x-direction, but I would like variability in the y-direction to better understand the variables that are important for a specific case.

```{r fig.height = 8, fig.width = 10}
hamby224_sensitivity_joined %>%
  filter(set == "Set 1") %>%
  droplevels() %>%
  select(situation, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(situation, case, rep, desc(feature_weight_abs)) %>%
  group_by(situation, case, rep) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(situation, case) %>%
  mutate(nlevels = length(levels(factor(feature)))) %>%
  ungroup() %>%
  arrange(situation, desc(nlevels), case) %>%
  mutate(order = rep(rep(1:length(levels(case)), each = 10), 22)) %>%
  ggplot(aes(x = order, fill = feature)) + 
  geom_bar(position = "stack", width = 1) + 
  coord_flip() + 
  facet_wrap(situation ~ ., ncol = 5) +
  #facet_grid(nlevels ~ situation) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  labs(x = "cases")

hamby224_sensitivity_joined %>%
  filter(set == "Set 1") %>%
  droplevels() %>%
  select(situation, case, rep, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(situation, case, rep, desc(feature_weight_abs)) %>%
  group_by(situation, case, rep) %>%
  slice(2) %>%
  ungroup() %>%
  group_by(situation, case) %>%
  mutate(nlevels = length(levels(factor(feature)))) %>%
  ungroup() %>%
  arrange(situation, desc(nlevels), case) %>%
  mutate(order = rep(rep(1:length(levels(case)), each = 10), 22)) %>%
  ggplot(aes(x = order, fill = feature)) + 
  geom_bar(position = "stack", width = 1) + 
  coord_flip() + 
  facet_wrap(situation ~ ., ncol = 5) +
  #facet_grid(nlevels ~ situation) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  labs(x = "cases")

>>>>

---
title: "03-computation_issues"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE)
```

@@

This journal documents some of the computation issues I have run into and my "solutions" or "conclusions".

```{r}
# Load libraries
library(tidyverse)
library(plotly)
library(caret)
library(randomForest)
library(lime)
```

# Plotly Dead Space Issue

@@

# Caret vs RandomForest

@@

# Comparing Furrr Times

@@

# Session Info

```{r}
sessionInfo()
```
